Cluster-efficient pangenome graph construction with nf-core/pangenome

### 2024-10-08 manuscript/main.tex
preparing the final number

	To compensate for that, we wrote \textit{nf-core/pangenome}, a reference-unbiased approach to construct pangenome graphs. 
	Mirroring PGGB, nf-core/pangenome is implemented in Nextflow \citep{DiTommaso2017}.
	% and follows the community-curated nf-core \citep{Ewels2020} best practice development guidelines. 
	%Providing all software dependencies in biocontainers \citep{daVeigaLeprevost2017} makes the pipeline portable and easy to install on HPC environments. 
	In contrast to PGGB, nf-core/pangenome can distribute the quadratic all-to-all base-level alignments across nodes of a cluster by splitting the approximate alignments into problems of equal size. 
	We benchmarked the time spent on base-pair level alignments and show that it is reduced linearly with an increase in alignment problem chunks (Suppl. ~\ref{aln_jobs_dist}). 
	We showcase the workflow’s scalability by applying it to 1000 chromosome 19 human haplotypes, and to 2146 \textit{E. coli} sequences, which were built in less than half the time PGGB required while not increasing the CO2 equivalent (CO2e) emissions.
	\vspace{-0.6cm}
	\section{Material and Methods}

@@

	\red{The Human Pangenome Resource Consortium (HPRC) recently built a draft human pangenome of 90 haplotypes \citep{Liao2023}. 
	However, haplotype data for thousands of individuals already exists, generated by the 1000 Genomes Project (1KGP) \citep{Durbin2010}. 
	As a use case, we used nf-core/pangenome to build a graph of 1000 chromosome 19 haplotypes \citep{Kuhnle2020} in 3 days, emitting 22.52 kg CO2e.
	PGGB took 7 days for the same task. 
	In Fig.~\ref{fig:hist_chr19} the pangenome growth curve generated with PANACUS \citep{Liao2023} shows nucleotide growth as more haplotypes are added.}
	\red{The softcore pangenome, defined as sequences traversed by 95\% of haplotypes, comprises the majority of the pangenome even with 1000 haplotypes. 
	This stability may be due to the exclusion of complex regions like the centromere in the short-read data.
	}

@@

	\red{Although} we expect our pipeline to scale for future challenges, such as for the next HPRC phase which targets 350 individuals\red{, further optimizations are possible}: 
	IMplicit Pangenome Graph (IMPG) (\href{https://github.com/ekg/impg}{https://github.com/ekg/impg}) which is a tool that extracts homologous loci from all genomes mapped to a specific target region. 
	This would allow us to break the whole genome multiple alignments into smaller pieces, construct a pangenome graph for each piece, and lace these together into a full graph with \href{https://github.com/pangenome/gfalace}{https://github.com/pangenome/gfalace}. 
	%Our first trials 
	%(\href{https://hackmd.io/vh0Ap65uQTm40DdleA-Oow}{https://hackmd.io/vh0Ap65uQTm40DdleA-Oow}) 
	%suggest following this direction.

@@

	Code and data resources for this manuscript and its figures are available in the public repository: \href{https://github.com/subwaystation/pangenome-paper}{https://github.com/subwaystation/pangenome-paper}.
	\vspace{-0.6cm}
	\section*{Acknowledgments}

>>>>

	To compensate for that, we wrote \textit{nf-core/pangenome}, a reference-unbiased approach to construct pangenome graphs. 
	Mirroring PGGB, nf-core/pangenome is implemented in Nextflow \citep{DiTommaso2017}.
	% and follows the community-curated nf-core \citep{Ewels2020} best practice development guidelines. 
	%Providing all software dependencies in biocontainers \citep{daVeigaLeprevost2017} makes the pipeline portable and easy to install on HPC environments. 
	In contrast to PGGB, nf-core/pangenome can distribute the quadratic all-to-all base-level alignments across nodes of a cluster by splitting the approximate alignments into problems of equal size. 
	We benchmarked the time spent on base-pair level alignments and show that it is reduced linearly with an increase in alignment problem chunks (Suppl. ~\ref{aln_jobs_dist}). 
	We showcase the workflow’s scalability by applying it to 1000 chromosome 19 human haplotypes and 2146 \textit{E. coli} sequences, which were built in less than half the time PGGB required while not increasing the CO2 equivalent (CO2e) emissions.
	\vspace{-0.6cm}
	\section{Material and Methods}

@@

	\red{The Human Pangenome Resource Consortium (HPRC) recently built a draft human pangenome of 90 haplotypes.
		% \citep{Liao2023}. 
	However, haplotype data for thousands of individuals already exists, generated by the 1000 Genomes Project (1KGP) \citep{Durbin2010}. 
	As a use case, we used nf-core/pangenome to build a graph of 1000 chromosome 19 haplotypes \citep{Kuhnle2020} in 3 days, emitting 22.52 kg CO2e.
	PGGB took 7 days for the same task (??.?? kg CO2e). 
	In Fig.~\ref{fig:hist_chr19} the pangenome growth curve generated with PANACUS \citep{Liao2023} shows nucleotide growth as more haplotypes are added.}
	\red{The softcore pangenome, defined as sequences traversed by 95\% of haplotypes, comprises the majority of the pangenome even with 1000 haplotypes. 
	This stability may be due to the exclusion of complex regions like the centromere in the short-read data.
	}

@@

	\red{Although} we expect our pipeline to scale for future challenges, such as for the next HPRC phase which targets 350 individuals\red{, further optimizations are possible}: 
	\red{The IMplicit Pangenome Graph (IMPG) (\href{https://github.com/ekg/impg}{https://github.com/ekg/impg}) tool extracts homologous loci from genomes mapped to a specific target region.}
	This would allow us to break the whole genome multiple alignments into smaller pieces, construct a pangenome graph for each piece, and lace these together into a full graph with \href{https://github.com/pangenome/gfalace}{https://github.com/pangenome/gfalace}. 
	%Our first trials 
	%(\href{https://hackmd.io/vh0Ap65uQTm40DdleA-Oow}{https://hackmd.io/vh0Ap65uQTm40DdleA-Oow}) 
	%suggest following this direction.

@@

	\red{Code and data resources for this manuscript and its figures are available in the public repository:} \href{https://github.com/subwaystation/pangenome-paper}{https://github.com/subwaystation/pangenome-paper}.
	\vspace{-0.6cm}
	\section*{Acknowledgments}
